High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-05-02 notebooks/gmsc.qmd
added code for real-world data experiments

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive, mod_name="model")
    conf_model = conformal_model(model; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    @info "Begin training $mod_name."
    fit!(mach)
    @info "Finished training $mod_name."
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
if _retrain
    model_dict = Dict(mod_name => _train(mod; mod_name=mod_name) for (mod_name, mod) in models)
    Serialization.serialize(joinpath(output_path,"cal_housing_models.jls"), model_dict)
else
    model_dict = Serialization.deserialize(joinpath(output_path,"cal_housing_models.jls"))
end
```

@@

# measure = Dict(
#     :f1score => multiclass_f1score, 
#     :acc => accuracy, 
#     :precision => multiclass_precision
# )
# model_performance = DataFrame()
# for (mod_name, mod) in model_dict
#     # Test performance:
#     test_data = load_mnist_test()
#     _perf = CounterfactualExplanations.Models.model_evaluation(mod, test_data, measure=collect(values(measure)))
#     _perf = DataFrame([[p] for p in _perf], collect(keys(measure)))
#     _perf.mod_name .= mod_name
#     model_performance = vcat(model_performance, _perf)
# end
# Serialization.serialize(joinpath(output_path,"cal_housing_model_performance.jls"), model_performance)
# CSV.write(joinpath(output_path, "cal_housing_model_performance.csv"), model_performance)
# model_performance
```

@@

bmk = benchmark(
    counterfactual_data; 
    models=model_dict, 
    generators=generator_dict, 
    measure=measures,
    suppress_training=true, dataname="California Housing",
    n_individuals=5,
    factual=0, target=1,
    initialization=:identity,
)
CSV.write(joinpath(output_path, "cal_housing_benchmark.csv"), bmk())
```

@@

```{julia}
df = @chain bmk() begin
    @filter(variable in [
        "distance_from_energy",
        "distance_from_targets",
        "distance",])
    @mutate(variable = ifelse.(variable .== "distance_from_energy", "Non-Conformity", variable))
    @mutate(variable = ifelse.(variable .== "distance_from_targets", "Implausibility", variable))
    @mutate(variable = ifelse.(variable .== "distance", "Cost", variable))
end
plt = AlgebraOfGraphics.data(df) * visual(BoxPlot) * 
    mapping(:generator, :value, row=:variable, col=:model, color=:generator)
plt = draw(
    plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=150, height=120), 
    facet=(; linkyaxes=:minimal)
)   
display(plt)
save(joinpath(output_images_path, "cal_housing_benchmark.png"), plt, px_per_unit=5)
```

>>>>

```{julia}
# Train models:
function _train(model, X=X, y=labels; cov=.95, method=:simple_inductive, mod_name="model")
    conf_model = conformal_model(model; method=method, coverage=cov)
    mach = machine(conf_model, X, y)
    @info "Begin training $mod_name."
    fit!(mach)
    @info "Finished training $mod_name."
    M = ECCCo.ConformalModel(mach.model, mach.fitresult)
    return M
end
if _retrain
    model_dict = Dict(mod_name => _train(mod; mod_name=mod_name) for (mod_name, mod) in models)
    Serialization.serialize(joinpath(output_path,"gmsc_models.jls"), model_dict)
else
    model_dict = Serialization.deserialize(joinpath(output_path,"gmsc_models.jls"))
end
```

@@

# measure = Dict(
#     :f1score => multiclass_f1score, 
#     :acc => accuracy, 
#     :precision => multiclass_precision
# )
# model_performance = DataFrame()
# for (mod_name, mod) in model_dict
#     # Test performance:
#     test_data = load_mnist_test()
#     _perf = CounterfactualExplanations.Models.model_evaluation(mod, test_data, measure=collect(values(measure)))
#     _perf = DataFrame([[p] for p in _perf], collect(keys(measure)))
#     _perf.mod_name .= mod_name
#     model_performance = vcat(model_performance, _perf)
# end
# Serialization.serialize(joinpath(output_path,"gmsc_model_performance.jls"), model_performance)
# CSV.write(joinpath(output_path, "gmsc_model_performance.csv"), model_performance)
# model_performance
```

@@

bmk = benchmark(
    counterfactual_data; 
    models=model_dict, 
    generators=generator_dict, 
    measure=measures,
    suppress_training=true, dataname="Californian Housing",
    n_individuals=5,
    factual=0, target=1,
    initialization=:identity,
)
CSV.write(joinpath(output_path, "gmsc_benchmark.csv"), bmk())
```

@@

```{julia}
df = @chain bmk() begin
    @filter(variable in [
        "distance_from_energy",
        "distance_from_targets",
        "distance",])
    @mutate(variable = ifelse.(variable .== "distance_from_energy", "Non-Conformity", variable))
    @mutate(variable = ifelse.(variable .== "distance_from_targets", "Implausibility", variable))
    @mutate(variable = ifelse.(variable .== "distance", "Cost", variable))
end
plt = AlgebraOfGraphics.data(df) * visual(BoxPlot) * 
    mapping(:generator, :value, row=:variable, col=:model, color=:generator)
plt = draw(
    plt, axis=(xlabel="", xticksvisible=false, xticklabelsvisible=false, width=150, height=120), 
    facet=(; linkyaxes=:minimal)
)   
display(plt)
save(joinpath(output_images_path, "gmsc_benchmark.png"), plt, px_per_unit=5)
```
